%==============================================================================
% Section 1 · Introduction (~1 page)
%
% Narrative arc (NeurIPS 2026 标准三段式):
%   Gap:  diffusion PDE solvers fail at shocks — empirically, analytically, algorithmically
%   Insight:  Score Shocks reveals the score is a Burgers field → double-Burgers coupling
%   Contribution:  (C1) coupling theorem / (C2) tanh-hard-coded architecture / (C3) ε^{1/2} rate
%
% 写作原则: 每段结尾指向下一段 / 每句话有信息增量 / 不用"I/We/Our" (双盲匿名)
%==============================================================================
\section{Introduction}
\label{sec:intro}

% ----- Gap: the shock triple-blind-spot -----
Generative diffusion in function space is reshaping data-driven {PDE} solving. The current frontier---{DiffusionPDE} \citep{huang2024diffusionpde} and {FunDPS} \citep{yao2025fundps} in Banach spaces---achieves striking accuracy on Darcy, Helmholtz, and moderate-Reynolds Navier--Stokes. Yet hyperbolic conservation laws remain untouched. The avoidance is methodological, not accidental: this class is the simultaneous \emph{empirical}, \emph{analytical}, and \emph{algorithmic} blind spot of the paradigm.

\textit{Empirically}, the five {PDE} benchmarks of \citet{yao2025fundps} are every one elliptic or parabolic; the sole Burgers experiment of \citet{huang2024diffusionpde} fixes $\physvis = 10^{-2}$ to sidestep the inviscid limit. \textit{Analytically}, existing convergence theory for diffusion solvers operates in $L^{2}$, $\Wass{2}$, or $\TV$ under classical smoothness; none accommodates weak solutions, $L^{1}$-contraction, or the Kruzhkov entropy condition \citep{kruzhkov1970}. \textit{Algorithmically}, the {PDE}-residual guidance central to both {DiffusionPDE} and {FunDPS} evaluates spatial derivatives by central differences---a discretisation that is singular at discontinuity interfaces. The guidance velocity itself diverges at the very feature that defines the hyperbolic regime.

% ----- Insight: the double-Burgers pivot -----
A recent identity changes the calculus. \citet{sarkar2026scoreshocks} establish that the score field of a {VE-SDE} diffusion model satisfies a viscous Burgers equation along the diffusion-time axis: $\dtau \score + 2\score\cdot\nabla_{u}\score = \Lap_{u}\score$. Equivalently, $-2\score$ is itself a Burgers velocity field. When the target distribution consists of solutions to a hyperbolic conservation law---itself of Burgers type in space-time---two Burgers structures coexist in orthogonal directions, and they are \emph{coupled}: physical shocks define the singular support of the data law, thereby fixing the location of the score-level interfacial layer through the correspondence $\Shockphys(t) \leftrightarrow \Shockscore(\tau)$ (Figure~\ref{fig:double-burgers}, Theorem~\ref{thm:double-burgers}).

This coupling reframes shock geometry from an obstacle into an asset. In standard diffusion training, a neural denoiser must \emph{learn} the sharp $\tanh$ profile of the score near a shock from finite data; the approximation error there is amplified by an $\exp(\amp T_{d})$ factor \citep[Theorem~6.3]{sarkar2026scoreshocks}, rendering the shock region fragile. The double-Burgers picture suggests a different route: \emph{hard-code} the exact $\tanh$ interfacial profile into the network architecture, reducing the learning task from reproducing a singular function to merely locating it and estimating its amplitude.

% ----- Contributions -----
The following three contributions execute this programme and provide its theoretical certification (quantitative validation in \S\ref{sec:experiments}):

\begin{enumerate}[label=(\textbf{C\arabic*}), leftmargin=*, itemsep=2pt]
    \item \textbf{Double-Burgers coupling (Theorem~\ref{thm:double-burgers}).} We formalise the geometric correspondence between $\Shockphys(t)$ and $\Shockscore(\tau)$ as a coupled Burgers system on $(\tau, t)$, establishing the structural foundation.
    \item \textbf{BV-aware score parameterization (\S\ref{sec:method}).} The exact $\tanh$ interfacial profile of the score field \citep[Prop.~5.4]{sarkar2026scoreshocks} is embedded into the network as an architectural prior: $\sth = \nabla\phisbg + (\jumpamp/2)\,\tanh(\phissh/2\sigtau^{2})\,\nabla\phissh$. The network learns only the shock location ($\phissh$) and amplitude ($\jumpamp$), not the shock shape.
    \item \textbf{Improved rate $\Wass{1} \le \mathcal{O}(\varepsilon^{1/2})$ (Theorem~\ref{thm:improved-rate}).} Under (C2) and a $\BV$ regulariser, generated samples converge to the Kruzhkov entropy solution at rate $\varepsilon^{1/2}$ in the score-matching error, with constants \emph{independent} of the Score Shocks amplification exponent $\amp$.
\end{enumerate}
