%==============================================================================
% Section 6 · Conclusion & Limitations (~0.5 page)
%
% NeurIPS 喜欢明确的 limitation 段。我们诚实标注：
%   - 理论目前只覆盖 1D 标量守恒律；多维 / 系统未证
%   - VE-SDE 假设；VP 留 future work
%   - shock 位置网络 \phi^{sh} 在 2D 上的工程难度
%==============================================================================
\section{Conclusion and Limitations}
\label{sec:conclusion}

% ----- Conclusion -----
\todo{Conclusion paragraph: We turned the Score Shocks observation — that a diffusion model's score itself satisfies a Burgers equation — from a purely theoretical curiosity into an algorithmic and theoretical lever for solving hyperbolic PDEs. The double-Burgers structure (Theorem~\ref{thm:double-burgers}) leads naturally to a BV-aware score parameterization, which (Theorem~\ref{thm:improved-rate}) achieves $\Wass{1} \le \mathcal O(\varepsilon^{1/2})$ convergence to the Kruzhkov entropy solution, free of the $\exp(\amp)$ blow-up of the standard analysis.}

% ----- Limitations -----
\paragraph{Limitations.}
\todo{Three explicit limitations:}
\begin{itemize}
    \item \textbf{Scalar 1D only (theory).} Theorems \ref{thm:improved-rate}–\ref{thm:shock-location} are stated for 1D scalar conservation laws. The system case (Euler) is supported empirically (E3) but a formal extension is left as future work.
    \item \textbf{VE-SDE only.} We work in the variance-exploding setting where Score Shocks gives a clean Burgers identity. The VP / EDM-Karras case is similar in spirit but the algebra is messier; we leave it to a follow-up.
    \item \textbf{$\phi^{\mathrm{sh}}$ in 2D.} Constructing a signed-distance-to-shock function in two dimensions is non-trivial; we use a mesh-based level-set surrogate in the appendix. A learned, mesh-free $\phi^{\mathrm{sh}}$ is an interesting follow-up.
\end{itemize}

% ----- Broader impact / future work -----
\paragraph{Future work.}
\todo{Three future directions (one sentence each): (i) extension to systems and multi-D; (ii) plasma kinetic equations (Vlasov–Poisson) leveraging the entropy-aware machinery; (iii) connection to flow-matching variants and Hilbert-space diffusion (FunDPS \citep{yao2025fundps}).}
