%==============================================================================
% Appendix A3 · Implementation Details
%==============================================================================
\section{Implementation Details}
\label{appx:impl}

\subsection{Network architecture}
\label{appx:impl:arch}

All experiments use a 1D U-Net backbone with four down/up-sampling stages and
GroupNorm (8 groups). The baseline and StandardScore variants use base
dimension $\dim=64$ ($\approx 0.4$\,M parameters); the BV-aware variant
(Eq.~\ref{eq:bv-aware}) uses $\dim=128$ ($\approx 7.7$\,M parameters). The
architecture follows the {EDM} preconditioning (Karras et al., 2022):
$D_{\theta}(x,\sigma) = c_{\mathrm{skip}}(\sigma)\,x + c_{\mathrm{out}}(\sigma)\,F_{\theta}(c_{\mathrm{in}}(\sigma)\,x,\;c_{\mathrm{noise}}(\sigma))$,
with $\sigma_{\mathrm{data}}=0.5$. The BV-aware parameterization decomposes
the score into three sub-networks: $\phisbg$ (the U-Net described above),
$\phissh$ (two Conv1d layers, 32 channels, kernel size 3, SiLU activation),
and $\jumpamp$ (two Conv1d layers, 16 channels, kernel size 1, SiLU, with a
Softplus gate enforcing $\jumpamp>0$). The Tweedie inversion
$D_{x} = x + \sigma^{2}s_{\theta}$ is applied at the output to maintain
compatibility with the EDM loss and sampler.

\subsection{Training}
\label{appx:impl:train}

All models are trained for the stated number of epochs (200 for BV-aware,
50 for StandardScore and Baseline) using the Adam optimiser with learning
rate $2\times10^{-4}$, batch size 64, no weight decay, and a fixed random
seed (PyTorch default; $\mathrm{seed}=0$ for data generation). The
viscosity-matched schedule sets $\physvis=1.0$, $T_{d}=1.0$, giving
$\sigma(\tau) = \sqrt{2\tau}$ with $\sigma\in[0,\sqrt{2}]$. The EDM baseline
uses the standard log-normal schedule $\ln\sigma\sim\mathcal{N}(-1.2,\,1.44)$.
Loss weights are $\lambda_{\mathrm{DSM}}=1.0$, $\lambda_{\BV}=0.1$ for all
EntroDiff configurations; the time-consistency term (Section~\ref{sec:exp:ablation})
uses $\lambda_{\mathrm{time}}=1.0$. Training was performed on a single NVIDIA
GeForce GPU with 24\,GB memory (RTX 3090/4090 class); one full run of the
BV-aware configuration (200 epochs) completes in approximately 35 minutes.

\subsection{Sampling}
\label{appx:impl:sample}

The Heun second-order integrator (Algorithm~\ref{alg:entrodiff-sample}) uses
$N_{\tau}=50$ steps as default, with ablations at $N_{\tau}\in\{10,25,50,100\}$.
The Godunov-form {PDE} guidance coefficient $\zeta_{\mathrm{PDE}}$ is set to
$0$ for unconditional generation; an ablation on $\zeta_{\mathrm{PDE}}\in\{0,0.1,0.5,1.0\}$
(Figure~A3 in Appendix~\ref{appx:extra-exp}) confirms that the guidance term is numerically
stable across all tested values but, for unconditional sampling, does not
improve $\Wass{1}$ beyond the architectural prior, since the $\Lpde^{\mathrm{Godunov}}$
minimiser at unconstrained initial conditions is the trivial solution $u\equiv0$.
The time-step of the Godunov residual matches the data generation:
$\Delta t=5\times10^{-3}$, $\Delta x=2\pi/128$.
The guidance gradient is clipped to unit norm per spatial channel to prevent
overflow from the large Godunov residuals of random initial noise.

\subsection{Compute and reproducibility}
\label{appx:impl:compute}

All experiments were conducted on a Linux workstation with three NVIDIA GPUs
(24\,GB each), 125\,GB RAM, and an NVMe SSD. Data generation (5,000 trajectories
of 1D Burgers at $N_{x}=128$) completes in under two minutes. The full
experimental pipeline, including data generation, training, and evaluation
scripts, is available in the anonymised supplementary material under the
MIT license. Key hyperparameters are summarised in Table~\ref{tab:hyperparams}.

\begin{table}[h]
    \caption{Hyperparameter summary.}
    \label{tab:hyperparams}
    \centering
    \begin{tabular}{ll}
        \toprule
        Parameter & Value \\
        \midrule
        Optimiser & Adam \\
        Learning rate & $2\times10^{-4}$ \\
        Batch size & 64 \\
        $\physvis$ (schedule) & 1.0 \\
        $T_{d}$ & 1.0 \\
        $\lambda_{\BV}$ & 0.1 \\
        Heun steps $N_{\tau}$ & 50 \\
        Data samples & 5,000 \\
        Spatial grid $N_{x}$ & 128 \\
        Temporal steps $N_{t}$ & 100 \\
        Physical time $T$ & 0.5 \\
        \bottomrule
    \end{tabular}
\end{table}
